Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan

### 2019-06-17 analysis/paper/paper.Rmd
added some text and one spatial plot

```{r tidy-data-for-table}
# clean some 
ornaments_period_clean <- 
ornaments_period %>% 
  mutate(Type = str_remove(Type, "prism/"),
         Type = ifelse(Type == "glass bead", "medium bead", Type),
         Type = ifelse(Type == "small I", "plain small", Type),
         Type = ifelse(Type == "small II", "thin small", Type))

@@

# Results

## Frequency between time periods

The results of the comparison between different time periods show a general pattern that most ornaments appeared in the pre-European contact period and the frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden beads. This trend can be also seen on other ornaments including agate bead, metal ring, and bell. However, glass bead shows a different pattern that indicates a higher frequency in the pre-European contact, and then the frequency decreased in the European contact period and even less in the Chinese contact period.

The distribution of frequency for subtypes in each major category are presented in (make a figure to combine all). Agate beads can be divided into 7 subtypes, in which hexagonal shape is the most common subtype that appeared before European contact and the number increased in the Eruopean contact and then declined in the Chinese contact. Small oval type shows similar frequency before and during European contact but not found in the Chinese contact period, while waxy oval bead only found during European contact period. There are more subtypes of agate beads during European contact compared to other time periods. For golden bead, we focus on their length

## Frequency for spatial pattern
2. Ornaments by: 
 - time period

@@

```{r agate-bead-fre-plot}
# plot for different shape of agate bead by period
agate_bead_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Agate bead')

@@

```{r metal-fre-plot}
# plot for different shape of metal ring by period
ring_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Metal ring')

@@

```{r}
library(cowplot)
plot_grid(agate_bead_shape_plot, ring_shape_plot,
          labels = c("A", "B"))

plot_grid(bell_shape_plot, glass_bead_plot,
          labels = c("C", "D"))
```

```{r scatter-plot-metal-by-period}
# plot for intact bell by period
ornaments_intact_bell <-
  ornaments_period %>% 
  filter(Condition %in% c('complete', 'near-complete'),
         Categories == 'Bell',
         Type != 'unclassified') %>% 
  mutate(Length = as.numeric(gsub("\\+", "", `Length(mm)`))) #remove +

ggplot(ornaments_intact_bell) +
  geom_point(aes(Length, Width, col = Type)) +
  facet_wrap(~ period) +
  labs(x = "Length (mm)",
       y = "Width (mm)",
       title = "Bell") +
  coord_fixed(ratio = 1) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

# plot for intact ring by period
ornaments_intact_ring <-
  ornaments_period %>% 
  filter(Condition %in% c('complete', 'near-complete'),
         Categories == 'Metal ring') %>% 
  mutate(Length = as.numeric(gsub("\\+|\\-", "", `Length(mm)`))) #remove + -

ggplot(ornaments_intact_ring) +
  geom_point(aes(Length, Width, col = Type)) +
  facet_wrap(~ period) +
  labs(x = "Length (mm)",
       y = "Width (mm)",
       title = "Metal ring") +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))
```

```{r read-in-the-spatial-data, eval = TRUE}

library(sf)
library(cartography)
burial <- st_read(here("analysis", "data", "raw_data", "AD_burial.shp"))
AD <- st_read(here("analysis", "data", "raw_data", "AD_zone.shp"))
AD_data <- st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"))
location <- st_read(here("analysis", "data", "raw_data", "location.shp"))
units_40 <- read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv"))
```

```{r tidy-data2, eval = TRUE}
# 40 sampling units
sample_units <-
  AD_data %>% 
  semi_join(units_40, by = c (`坑號` = "the_sq"))

@@

@@

```{r-gganimate}
# test gganimate for measurement of beads by period
# devtools::install_github('thomasp85/gganimate')
library(gganimate)

@@

>>>>

```{r tidy-data-for-table}
# clean some 
ornaments_period_clean <- 
ornaments_period %>% 
  mutate(Type = str_remove(Type, "prism/"),
         Type = ifelse(Type == "glass bead", "medium bead", Type),
         Type = ifelse(Type == "small I", "plain small", Type),
         Type = ifelse(Type == "small II", "thin small", Type),
         Type = ifelse(Type == "large thin string", "thin large", Type))

@@

@@

```{r agate-bead-fre-plot}
# plot for different shape of agate bead by period
agate_bead_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Agate bead',
         Type != 'other')

@@

```{r metal-fre-plot}
# plot for different shape of metal ring by period
ring_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Metal ring',
         !Type %in% c('large thick string',
                       'small thin string'))

@@

```{r plot-altogether, fig.cap= "Frequency of "}
library(cowplot)
plot_grid(agate_bead_shape_plot, ring_shape_plot,
          bell_shape_plot, glass_bead_plot,
          labels = c("A", "B", "C", "D"), ncol=2)
```
# Results

## Frequency between time periods
The results of the comparison between different time periods show a general pattern that most ornaments appeared in the pre-European contact period and the frequencies reached a peak during the European contact and then dropped during the Chinese contact period, especially for golden beads. This trend can be also seen on other ornaments including agate bead, metal ring, and bell. However, glass bead shows a different pattern that indicates a higher frequency in the pre-European contact, and then the frequency decreased in the European contact period and even less in the Chinese contact period.

The distribution of frequency for subtypes in each major category are presented in \@ref(fig:plot-all-ornaments-count). The distribution shows that agate beads and metal rings have greater quantity and variety of shapes compared to copper bell and glass beads during the European contact period. Agate beads can be divided into eight subtypes, in which hexagonal shape is the most common subtype that appeared before European contact and increased during the European contact and then declined in the Chinese contact. Small oval type shows similar frequency before and during European contact but not found in the Chinese contact period, while waxy oval bead only found during European contact period. There are more subtypes of agate beads during European contact compared to other time periods. Similarly metal rings can be classified into seven subtypes, in which wide small ring is the most common metal ornaments, followed by wide large and thin large shape. The greater varieties for those two categories might indicate the multiple sources due to global trade network brought by the Europeans. In contrast, copper bell and glass bead have smaller variety, but glass bead has larger number due to they were usually found in a cluster that reflect the original use as strings. Most common shape for copper bell is the large shape with human face as motif, while most common shape for glass bead is small that were identified as Indo-Pacific beads, the most common type that are widespread in Southeast Asian sites since 300 BC [@Francis2002].

## Frequency for spatial pattern
Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of the ornaments for each time period. It shows that ornaments concentrate on one specific square in the pre-European contact, and more ubiquitous during the European contact and

```{r scatter-plot-metal-by-period, eval= FALSE}
# plot for intact bell by period
ornaments_intact_bell <-
  ornaments_period %>% 
  filter(Condition %in% c('complete', 'near-complete'),
         Categories == 'Bell',
         Type != 'unclassified') %>% 
  mutate(Length = as.numeric(gsub("\\+", "", `Length(mm)`))) #remove +

ggplot(ornaments_intact_bell) +
  geom_point(aes(Length, Width, col = Type)) +
  facet_wrap(~ period) +
  labs(x = "Length (mm)",
       y = "Width (mm)",
       title = "Bell") +
  coord_fixed(ratio = 1) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))
```

```{r read-in-the-spatial-data, eval = TRUE}
library(sf)
library(cartography)
burial <- st_read(here("analysis", "data", "raw_data", "AD_burial.shp"))
AD <- st_read(here("analysis", "data", "raw_data", "AD_zone.shp"))
AD_data <- st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"))
location <- st_read(here("analysis", "data", "raw_data", "location.shp"))
units_40 <- read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv"))
```

```{r tidy-spatial-data, eval = TRUE}
# 40 sampling units
sample_units <-
  AD_data %>% 
  semi_join(units_40, by = c (`坑號` = "the_sq"))

@@

```{r plot-all-spatial-pattern, fig.cap= "Spatial pattern of ornaments"}
ggplot(ornaments_period_join) +
  geom_sf(data = sample_units, fill = NA) +
  geom_sf(data = ornaments_period_join, aes(fill =n), alpha = 0.9) +
  facet_wrap(~ period) +
  labs(fill = "frequency") +
  scale_fill_gradient(high = "black", low = "grey") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 11))
```

@@

```{r-gganimate, eval= FALSE}
# test gganimate for measurement of beads by period
# devtools::install_github('thomasp85/gganimate')
library(gganimate)

@@

Word count: `r wordcountaddin::word_count()`
